---
output: github_document
---

# aRlegislation

[![Build Status](https://travis-ci.com/titaniumtroop/aRlegislation.svg?branch=master)](https://travis-ci.com/titaniumtroop/aRlegislation)

This package provides a dataset of Arkansas legislation from 2001-2019 for data exploration, natural language processing, and graph analysis. The dataset is structured as a nested tibble containing distinct legislative sessions, nested by cycle (every two years) and session type. 

Because the package uses nested tibbles, the `tidyr` package is required to use the dataset:

```{r libraries, warning = F, message=F}
library(aRlegislation)
library(tidyr) # needed for nest/unnest operations
```

The structure of the dataset is:

```{r legislation_head}
head(legislation)
```

There are three different datasets for each legislative cycle and session:

* `acts`: contains act number, bill number, title, subtitle, text, and partisanship
* `lawmakers`: contains demographic information about elected senators and representatives such as chamber, political party, and district
* `sponsorship`: a simple table containing act number and sponsor name

The vignettes provide more detail about each type of nested tibble, and also provide an example or two of analyses that can be performed using each type of tibble.

# Future Plans

The legislature meets once a year, plus special sessions, so the dataset can be updated as each additional session concludes.

This dataset is essentially an aggregation of distinct datasets across each cycle/session of the legislature. This works well for seeing trends across cycles/sessions, but it doesn't give much visibility into how individual lawmakers act as they get re-elected, gain experience, and perhaps graduate from House to Senate as term limits force them out of office. As such, entity resolution of lawmakers across cycles/sessions will likely be the focus of the next major version of this dataset. 
