---
output: github_document
---

# Overview

This package provides a dataset of Arkansas legislation from 2001-2019 for data exploration, natural language processing, and graph analysis. The dataset is structured as a nested tibble containing distinct legislative sessions, nested by cycle (every two years) and session type. Before we can take a look, we'll need to load a couple of packages from the tidyverse:

```{r, warning = F, message=F}

library(aRlegislation)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyr)

```

Now, we can take a look at the structure of the dataset:

```{r}

head(legislation)

```

There are three different datasets for each legislative cycle and session:

* acts, which contains act number, bill number, title, subtitle, text, and partisanship
* sponsors, which contains demographic information about lawmakers such as chamber, political party, and district
* sponsorship, which is a simple table containing only act number and sponsor name in a manner suitable for creating network graphs

Below, we'll take a look at each type of nested tibble, and we'll provide an example or two of analyses that can be performed using each type of tibble.

## Sponsors

For purposes of this dataset, the term "sponsor" refers to one or more lawmakers that propose an enacted bill during a legislative session. While a sponsor can be an individual lawmaker or a committee, this table excludes committees. Here's what the structure of one of the sponsors tibbles looks like: 

```{r}

head(legislation$sponsors[[1]])

```

As an example, we can look at the political makeup of the lawmakers in each regular-session legislative chamber over time pretty easily:[^sponsorsNote]

[^sponsorsNote]: To keep things simple, we'll just look at regular legislative sessions for now. Also, committees have been excluded from this graph because they don't necessarily belong to a specific chamber.

```{r cycle-chamber-party, fig.width=7.5, fig.height=3}

# Use party-related colors
party.colors <- c(
  "R" = "#990000", # dark red = Republicans
  "D" = "#668cff", # light blue = Democrats
  "G" = "#00cc00", # bright green = Green party
  "I" = "#444444", # dark grey = Independent
  "unk" = "#b3b300", # dark yellow = unknown
  "bipartisan" = "#8A2BE2", # purple
  "committee" = "#888888" # medium grey
)

legislation %>%
  filter(grepl("R", session)) %>%
  unnest(sponsors) %>%
  filter(! grepl("Budget|Committee|Management|Senate|Efficiency|Insurance|Agencies", sponsor)) %>%
  count(cycle, session, party, chamber) %>%
  ggplot(aes(y = n, x = cycle, color = party)) +
    geom_point(size = 2) +
    geom_line(size = 1) +
    facet_wrap(~ chamber) +
    scale_x_continuous(
      breaks = seq(from = 2001, to = 2019, by = 4), 
      minor_breaks = seq(from = 2001, to = 2019, by = 2)
    ) +
    scale_color_manual(values = party.colors) +
    labs(
      x = "Legislative Cycle",
      y = "No. of Legislators",
      title = "Political Makeup of the Arkansas Legislature"
    )

```

This graph shows a very pronounced switch in the makeup of both chambers of the Arkansas legislature between 2009 and 2017. 

## Sponsorship

Sponsorship refers to individual laws that each lawmaker presents to the legislative chamber. Some bills have a single sponsor, while others have many. Bills can be sponsored by lawmakers in one or both chambers, or by a committee belonging to neither chamber. Since the number of lawmakers/committees who sponsor laws can change, this table is a two-column layout with a unique combination of act and sponsor per row:

```{r}

head(legislation$sponsorship[[1]])

```

The purpose of this simple layout is to facilitate graph analysis. The sponsor table contains additional information about vertices (where vertices are lawmakers), and the sponsorship table contains information about the edges.

A simple example from the 2019 regular session illustrates:

```{r}

### New tidygraph code

```


## Acts

The acts tibbles contain the text of laws passed during each cycle/session, along with a little bit of metadata about the acts: 

```{r}

head(legislation$acts[[1]])

```

The act text has been lightly processed such that it is still human-readable, especially in monospace type. A really short act provides an example: 

<blockquote>
`r legislation$acts[[1]]$text[[200]]`
</blockquote>

The metadata includes a calculated partisan value based on the political party of the bill sponsors, as follows:

* If any sponsor is a committee, the partisan value is set to committee
* If all sponsors are in the same party, the partisan value is set to the value of that party
* If the sponsors are of mixed political parties, the partisan value is set to bipartisan

The partisan value allows us to look at trends about bipartisanship over time, as follows:

```{r, fig.width=7.5, fig.height=3.5}

legislation %>%
  unnest(acts) %>%
  filter(partisan == "bipartisan") %>%
  count(cycle, partisan) %>% # counts number of parties sponsoring acts
  ggplot(aes(y = n, x = cycle, color = partisan)) +
    geom_point(size = 2.5) +
    geom_line(size = 3) +
    geom_smooth(method = "lm", color = "blue", se = T) +
    scale_x_continuous(
      breaks = seq(from = 2001, to = 2019, by = 4), 
      minor_breaks = seq(from = 2001, to = 2019, by = 2)
    ) +
    scale_color_manual(values = party.colors) +
    labs(
      title = "Total Bipartisan Acts",
      x = "",
      y = "",
      caption = ""
    ) +
    theme(
      legend.text = element_text(size = 6),
      legend.position = "none"
    )

```

We can see here that the total number of bipartisan acts has been declining over time.

# Future Plans

This dataset is essentially an aggregation of distinct datasets across each cycle/session of the legislature. This works well to allow us to see trends across cycles/sessions, but it doesn't give much visibility into how individual lawmakers act as they get re-elected, gain experience, and perhaps graduate from House to Senate as term limits force them out of office. As such, entity resolution of lawmakers across cycles/sessions will likely be the focus of the next major version of this dataset. 
