---
title: "Acts"
output:
  github_document
vignette: >
  %\VignetteIndexEntry{Acts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


This vignette examines the structure of the `acts` tibbles contained within the `legislation` dataset and displays the partisanship of the Arkansas Legislature over time as an exmample use case. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F, warning=F}
library(aRlegislation)
library(dplyr)
library(tidyr) # needed for nest/unnest operations
library(ggplot2)
```

The `acts` tibbles contain the text of laws passed during each cycle/session, along with a little bit of metadata about the acts: 

```{r acts_head}
head(legislation$acts[[1]])
```

The act text has been lightly processed such that it is still human-readable. A really short act provides an example: 

<blockquote>
`r legislation$acts[[1]]$text[[200]]`
</blockquote>

The act titles, subtitles, and text are suitable for natural language processing analysis, although examples of such are beyond the scope of this introduction. The text above illustrates one caveat: for acts that are amended, both the deleted text and the new text appears in the corpus [e.g., the act referenced above increases the dollar threshold from ~~ten thousand dollars ($10,000)~~ to __twenty thousand dollars ($20,000)__]. 

The metadata includes a calculated partisan value based on the political party of the bill sponsors, as follows:

* If any sponsor is a committee, the partisan value is set to committee
* If all sponsors are lawmakers in the same party, the partisan value is set to the value of that party
* If the sponsors are lawmakers of mixed political parties, the partisan value is set to bipartisan

The partisan value allows us to look at trends about bipartisanship over time, as follows:

```{r bipartisanship_plot, fig.width=7, fig.height=3.5}
party.colors <- c(
  "R" = "#990000", # dark red = Republicans
  "D" = "#668cff", # light blue = Democrats
  "G" = "#00cc00", # bright green = Green party
  "I" = "#444444", # dark grey = Independent
  "unk" = "#b3b300", # dark yellow = unknown
  "bipartisan" = "#8A2BE2", # purple
  "committee" = "#888888" # medium grey
)

legislation %>%
  unnest(acts) %>%
  filter(partisan == "bipartisan") %>%
  count(cycle, partisan) %>% # counts number of parties sponsoring acts
  ggplot(aes(y = n, x = cycle, color = partisan)) +
    geom_point(size = 2.5) +
    geom_line(size = 3) +
    geom_smooth(method = "lm", color = "blue", se = T) +
    scale_x_continuous(
      breaks = seq(from = 2001, to = 2019, by = 4), 
      minor_breaks = seq(from = 2001, to = 2019, by = 2)
    ) +
    scale_color_manual(values = party.colors) +
    labs(
      title = "Total Bipartisan Acts",
      x = "",
      y = "",
      caption = ""
    ) +
    theme(
      legend.text = element_text(size = 6),
      legend.position = "none"
    )
```

We can see here that the total number of bipartisan acts has been declining over time.

